\documentclass{article}

\usepackage[a4paper]{geometry}
\usepackage{hyperref}

\title{commutative.group}

\begin{document}
\maketitle
\newline
\begin{center}
  A Blog on Math, Algorithms, and Symmetry by Anakin Dey
\end{center}
\begin{table}
  \centering
  \begin{tabular}{cc}
    \LARGE[\href{About.html}{About}] & \LARGE[\href{All-Posts.html}{All}]
  \end{tabular}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{|l|}
    \hline
    \large 2024-01-07 \\
    \LARGE \href{posts/other_name.html}{A Group by Any Other Name} \\
    As is tradition, let $G$ be a group. But, can we define what a group is in more interesting ways? \\
    \hline
  \end{tabular}
\end{table}

\part*{\centering About}

\newline

My name is Anakin Dey.
I am a senior studying mathematics at the University of Illinois Urbana-Champaign.
As evidenced by this blog, my main interests in mathematics and computer science are on the algebraic and algorithmic side.
You can find more about me at \href{https://www.anakin-dey.com/}{my website}.

I have a work in progress tool I am calling \href{https://github.com/Spamakin/abel}{Abel} which generates this site from some \LaTeX\ files as well as some other templating.
The tool generates \href{https://github.com/Spamakin/commutative.group/}{code} which is then deployed to this site.
If you have any errors, you can email me at the email on my website or raise an issue on the Abel repository.
All blog posts are viewable not just as webpages, but also as PDFs if that is your thing.
Simply add \texttt{.pdf} to the url in the browser.

\part*{\centering All Posts}

\newline

Here is a list of all posts to date in reverse chronological order:

\begin{itemize}
\item 2024-01-07: \href{posts/other_name.html}{A Group by Any Other Name}
\end{itemize}

\part*{404}

\newline

Page not found!

\end{document}
